---
title: "Benchmarking 'arrangements'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking 'arrangements'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of this document is to compare the performances of some of the existing packages for generating combinations, permutations and partitions.

- [gtools](https://CRAN.R-project.org/package=gtools) 
    `r packageDescription("gtools")$Version`
- [combinat](https://CRAN.R-project.org/package=combinat) 
    `r packageDescription("combinat")$Version`
- [multicool](https://CRAN.R-project.org/package=multicool) 
    `r packageDescription("multicool")$Version`
- [partitions](https://CRAN.R-project.org/package=partitions)
    `r packageDescription("partitions")$Version`
- [permute](https://CRAN.R-project.org/package=permute)
    `r packageDescription("permute")$Version`
- [permutations](https://CRAN.R-project.org/package=permutations)
    `r packageDescription("permutations")$Version`
- [RcppAlgos](https://CRAN.R-project.org/package=RcppAlgos)
    `r packageDescription("RcppAlgos")$Version`

```{r, warning = FALSE, message = FALSE}
library(gtools)
library(combinat)
library(multicool)
library(partitions)
library(permute)
library(permutations)
library(RcppAlgos)
library(arrangements)
gtools.permutations <- gtools::permutations
gtools.combinations <- gtools::combinations
partitions.compositions <- partitions::compositions
utils.combn <- utils::combn
```

```{r, echo = FALSE}
library(microbenchmark)
library(ggplot2)
options(digits = 3)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

`gtools::permutations`, `gtools::combinations`, `partitions::compositions` and `utils::combn` are renamed because they are being overwritten by other packages.

### Permutations

**Permutations of 6 items**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
autoplot(microbenchmark(
    gtools = gtools.permutations(6, 6),
    multicool = allPerm(initMC(1:6)),
    partitions = partitions::perms(6),
    permuate = allPerms(6),
    permutations = allperms(6),
    RcppAlgos = permuteGeneral(6, 6),
    arrangements = permutations(6)
))
```

**5-Permutations of 7 items**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
autoplot(microbenchmark(
    gtools = gtools.permutations(7, 5),
    RcppAlgos = permuteGeneral(7, 5),
    arrangements = permutations(7, 5)
))
```

**Permutations of multiset**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r, echo = FALSE}
f <- c(2, 1, 2, 1, 2)
x <- rep(1:5, f)
autoplot(microbenchmark(
    multicool = allPerm(initMC(x)),
    RcppAlgos = permuteGeneral(5, freqs = f),
    arrangements = permutations(freq = f)
))
```

**K-Permutations of multiset**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
f <- c(2, 1, 2, 1, 2)
x <- rep(1:5, f)
autoplot(microbenchmark(
    multicool = allPerm(initMC(x)),
    RcppAlgos = permuteGeneral(5, 6, freqs = f),
    arrangements = permutations(k = 6, freq = f)
))
```

### Combinations

**7-Combinations of 15 items**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
autoplot(microbenchmark(
    gtools = gtools.combinations(15, 7),
    combinat = combn(15, 7),
    utils = utils.combn(15, 7),
    RcppAlgos = comboGeneral(15, 7),
    arrangements = combinations(15, 7)
))
```

**Combinations with replacement of 9 items**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
autoplot(microbenchmark(
    gtools = gtools.combinations(9, 6, repeats.allowed = TRUE),
    RcppAlgos = comboGeneral(9, 6, repetition = TRUE),
    arrangements = combinations(9, 6, replace = TRUE)
))
```

**Combinations of multiset**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
f <- c(2, 1, 2, 1, 2, 2, 1, 2)
x <- rep(1:8, f)
autoplot(microbenchmark(
    RcppAlgos = comboGeneral(8, 6, freqs = f),
    arrangements = combinations(k = 6, freq = f)
))
```

### Partitions


**Partitions of 30**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
autoplot(microbenchmark(
    partitions = parts(30),
    arrangements = partitions(30)
))
```

**Partitions of 50 into 10 parts**

```{r, echo = FALSE, results = "hide"}
gc()
```

```{r}
autoplot(microbenchmark(
    partitions = restrictedparts(50, 10, include.zero = FALSE),
    arrangements = partitions(50, 10)
))
```


**Compositions of 8**
```{r}
autoplot(microbenchmark(
    partitions = partitions.compositions(8, include.zero = FALSE),
    arrangements = compositions(8)
))
```

**Compositions of 20 into 5 parts**
```{r}
autoplot(microbenchmark(
    partitions = partitions.compositions(20, 5, include.zero = FALSE),
    arrangements = compositions(20, 5)
))
```
