---
title: "Benchmarking 'arrangements'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking 'arrangements'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of this document is to compare the performances of some of the existing packages for generating combinations, permutations and partitions.

- [gtools](https://CRAN.R-project.org/package=gtools) 
    `r packageDescription("gtools")$Version`
- [combinat](https://CRAN.R-project.org/package=combinat) 
    `r packageDescription("combinat")$Version`
- [multicool](https://CRAN.R-project.org/package=multicool) 
    `r packageDescription("multicool")$Version`
- [partitions](https://CRAN.R-project.org/package=partitions)
    `r packageDescription("partitions")$Version`
- [permute](https://CRAN.R-project.org/package=permute)
    `r packageDescription("permute")$Version`
- [permutations](https://CRAN.R-project.org/package=permutations)
    `r packageDescription("permutations")$Version`
- [RcppAlgos](https://CRAN.R-project.org/package=RcppAlgos)
    `r packageDescription("RcppAlgos")$Version`
- [iterpc](https://CRAN.R-project.org/package=iterpc): it is omitted as `arrangements` is thought of a replacement of `iterpc`.

```{r, echo = F}
suppressPackageStartupMessages({
    library(microbenchmark)
    library(gtools)
    library(combinat)
    library(multicool)
    library(partitions)
    library(permute)
    library(permutations)
    library(RcppAlgos)
    library(arrangements)
    gtools.permutations = gtools::permutations
    gtools.combinations = gtools::combinations
    ucombn = utils::combn
})
options(digits=3)
```

### Permutations

**Permutation of 7 items**

```{r}
microbenchmark(
    gtools = gtools.permutations(7, 7),
    multicool = allPerm(initMC(1:7)),
    partitions = partitions::perms(7),
    permuate = allPerms(7),
    permutations = allperms(7),
    RcppAlgos = permuteGeneral(7, 7),
    arrangements = permutations(7),
    times = 10,
    unit = "relative"
)
```

**5-Permutation of 7 items**

```{r}
microbenchmark(
    gtools = gtools.permutations(7, 5),
    RcppAlgos = permuteGeneral(7, 5),
    arrangements = permutations(7, 5),
    unit = "relative"
)
```

**Permutation of multiset**

```{r}
f <- c(2, 1, 2, 1, 2)
x <- rep(1:5, f)
microbenchmark(
    multicool = allPerm(initMC(x)),
    RcppAlgos = permuteGeneral(1:5, freqs = f),
    arrangements = permutations(f = f),
    unit = "relative"
)
```

**K-Permutation of multiset**

```{r}
f <- c(2, 1, 2, 1, 2)
x <- rep(1:5, f)
microbenchmark(
    multicool = allPerm(initMC(x)),
    RcppAlgos = permuteGeneral(5, 6, freqs = f),
    arrangements = permutations(f = f, k = 6),
    unit = "relative"
)
```

### Combinations

**7-Combinations of 15 items**

```{r}
microbenchmark(
    gtools = gtools.combinations(15, 7),
    combinat = combn(15, 7),
    utils = ucombn(15, 7),
    RcppAlgos = comboGeneral(15, 7),
    arrangements = combinations(15, 7),
    unit = "relative"
)
```

**Combinations with replacement of 9 items**

```{r}
microbenchmark(
    gtools = gtools.combinations(9, 6, repeats.allowed = TRUE),
    RcppAlgos = comboGeneral(9, 6, repetition = TRUE),
    arrangements = combinations(9, 6, replace = TRUE),
    unit = "relative"
)
```

**Combinations of multiset**

```{r}
f <- c(2, 1, 2, 1, 2, 2, 1, 2)
x <- rep(1:8, f)
microbenchmark(
    RcppAlgos = comboGeneral(8, 6, freqs = f),
    arrangements = combinations(f = f, k = 6),
    unit = "relative"
)
```

### Partitions


**Partitions of 30**

```{r}
microbenchmark(
    partitions = parts(30),
    arrangements = partitions(30),
    unit = "relative"
)
```

**Partitions of 50 into 10 parts**

```{r}
microbenchmark(
    partitions = restrictedparts(50, 10),
    arrangements = partitions(50, 10),
    unit = "relative"
)
```
